The Spine of American Legal Practice

### 2016-06-25 article/article.Rmd
Cite digital legal history

---
author:
- Kellen Funk
- Lincoln Mullen
fontsize: 12pt
output:
  pdf_document:
    fig_caption: True
    number_sections: False
    md_extensions: -tex_math_single_backslash
title: The Spine of American Legal Practice
---

@@

Because most codes were public statutes, they were widely printed and distributed and therefore found their way into libraries digitized by Google Books. Where possible, we collected our corpus from Google's publicly accessible database. We used optical character recognition software (OCR) to create plain-text versions of the codes, which we edited lightly---we corrected section markers by hand as necessary and wrote a script to fix the most common OCR errors.[^32]

The most important step we took in processing the files was to split each section of the code into its own text file. Codes varied in their organizations, but they universally divided specific regulations into *sections* (or, on occasion, *articles*). The discursive formation of these texts not only provides a handy organizational scheme for digital approaches, sections were historically the way legislators borrowed their texts.[^33] Codifiers took their sources apart by sections, rearranging here, editing, drafting, and then re-combining there. Despite the fact that states differed widely on what topics they included in "civil procedure," sectioning the codes allowed us to assess similarity even among codes of quite different lengths and coverage. For instance, we know from historical evidence that California's 1851 code was derived from New York's 1850 code. (Stephen Johnson Field, David Dudley's brother, was the lawyer who imported New York's code into California.[^34]) But the New York code is over 150,000 words long, where was California's code was just over 50,000 words long. Those rather disparate lengths mean that comparing all of the California code to all of the New York code is a less meaningful comparison that than comparing each section in the California code to each section in the New York code, where matching sections will be a similar length.

The next step was to compare sections to one another and measure the similarity between them. To continue the New York-to-California example, consider the following pairs of sections. The first pair is from the final draft of the New York Field Code. These sections completely abolish prior practice and begin to rebuild the procedure system from the ground up:

@@

A well-known method for measuring the similarity of two documents involves dividing texts up into "tokens" of consecutive words and calculating a Jaccard score, defined as ratio between the number of tokens that the two document have in common to the total number of tokens that appear in both documents. We used five-word tokens and "shingled" them, meaning that for the New York sections above, the first token was "the distinction between actions at," the second token was "distinction between actions at law," and so on. These tokens each contain more meaning than a single word, yet because they are shingled they are robust to changes in the text or noisy OCR. A Jaccard score (in set notation, EQUATION) will always produce a number between 0 (complete dissimilarity) and 1 (complete similarity).

@@

[^31]: We have preferred to use the term "digital history" when referring to our own work, in part because the term "digital humanities" has largely come to refer to the work of digital literary and digital media scholars, but primarily because we wish to see digital scholars make disciplinary, rather than interdisciplinary contributions. On the role of disciplines and the importance of field specific argumentation, see, Stephen Robertson, "The Differences between Digital Humanities and Digital History," {PAGE}--{PAGE}, and Cameron Blevins, "Digital History's Perpetual Future Tense," {PAGE}--{PAGE}, both in *Debates in the Digital Humanities 2016*, ed. Matthew K. Gold and Lauren F. Klein (Minneapolis: University of Minnesota Press, 2016). Nevertheless, digital text analysis in the humanities has mostly been published by literary scholars, including Stephen Ramsay, *Reading Machines: Toward an Algorithmic Criticism* (Urbana: University of Illinois Press, 2011); Franco Moretti, *Distant Reading* (New York: Verso, 2013); Matthew L. Jockers, *Macroanalysis: Digital Methods and Literary History* (Urbana: University of Illinois Press, 2013); Ted Underwood, *Why Literary Periods Mattered: Historical Contrast and the Prestige of English Studies* (Stanford: Stanford University Press, 2013). Related but less well regarded by humanities scholars is work in "culturomics": see Jean-Baptiste Michel et al., "Quantitative Analysis of Culture Using Millions of Digitized Books", *Science* 331, no. 6014 (2011), 176--182; doi:10.1126/science.1199644. On the uselessness of the term "big data," see Ted Underwood, "Against (Talking About) 'Big Data,'" *The Stone and the Shell*, blog post, 10 May 2013: <https://tedunderwood.com/2013/05/10/why-it-matters-that-we-dont-know-what-we-mean-by-big-data/>. For an example of forming a corpus, see Ted Underwood, Boris Capitanu, Peter Organisciak, Sayan Bhattacharyya, Loretta Auvil, Colleen Fallaw, J. Stephen Downie, "Word Frequencies in English-Language Literature, 1700-1922," dataset, v0.2 (HathiTrust Research Center, 2015) doi:10.13012/J8JW8BSJ.

[^32]: **Note codes pulled from HeinOnline and three codes transcribed by hand**.We favored public domain sources where available both for reasons of efficiency (large statute books download much more quickly from Google's servers than from subscription databases) and reproducibility (most of this project can be replicated without subscription databases). In each instance we downloaded an entire volume of sessions laws, statutory compilations, or single-volume codes of procedure and then cropped out irrelevant pages, marginalia and footnoted commentary, leaving only the statutory text. After several trials of various implementations of Tesseract (open source) and I.R.I.S. (proprietary) OCR programs, we determined that Nitro Pro PDF, which relies on I.R.I.S. software, offered the best OCR tool for this project. I.R.I.S. provides slightly more accurate readings of nineteenth-century typefaces than Tesseract, and Nitro Pro's implementation makes words, not characters, the fundamental unit of output. The latter feature made cropping between marginalia and the statute more reliable. We programmatically removed hyphenated line breaks and standardized spelling for common terms that evolved over the nineteenth century (e.g., *indorsement*).

[^33]: Cf. Hay & Craven, AJLH article

[^34]:

>>>>

---
author:
- Kellen Funk
- Lincoln Mullen
fontsize: 12pt
output:
  pdf_document:
    fig_caption: True
    md_extensions: "-tex_math_single_backslash"
    number_sections: False
title: The Spine of American Legal Practice
---

@@

Because most codes were public statutes, they were widely printed and distributed and therefore found their way into libraries digitized by Google Books. We drew primarily from the Google Books, filling in gaps from other databases as necessary.[^32] We used optical character recognition software (OCR) to create plain-text versions of the codes, which we edited lightly, correcting section markers by hand as necessary and writing a script to fix the most obvious OCR errors.[^33]

The most important step we took in processing the files was to split each section of the code into its own text file. Codes varied in how they were organized, but they universally divided specific regulations into *sections* (or, on occasion, *articles*). Not only does the discursive form of these texts provide a handy organizational scheme for digital methods, but historically sections were also the way legislators borrowed their texts. Codifiers took their sources apart by sections, rearranging here, editing, drafting, and then re-combining there. Despite the fact that states differed widely on what topics they included in "civil procedure," sectioning the codes allowed us to assess similarity even among codes of quite different lengths and coverage. For instance, we know that California's 1851 code was derived from New York's 1850 code. (Stephen Johnson Field, David Dudley Field's brother, was the lawyer who imported New York's code into California.[^34]) But the New York code is over 150,000 words long, whereas California's code was just over 50,000 words long. Those disparate lengths mean that comparing all of the California code to all of the New York code is a less meaningful comparison that than comparing each section in the California code to each section in the New York code, where matching sections will have a similar length.

Having divided the texts according to a historically justified pattern, our next step was to compare each section to every other section and measure the similarity between them. To continue the New York-to-California example, consider the following pairs of sections. The first pair is from the final draft of the New York Field Code. These sections completely abolished prior practice and began to rebuild the procedure system from the ground up:

@@

A well-known method for measuring the similarity of two documents involves dividing texts up into tokens of consecutive words (called n-grams) and calculating a Jaccard similarity score, defined as ratio between the number of tokens that the two document have in common to the total number of tokens that appear in both documents. We used five-word tokens and "shingled" them, meaning that for the New York sections above, the first token was "the distinction between actions at," the second token was "distinction between actions at law," and so on. These tokens each contain more meaning than a single word, yet because they are shingled they are robust to changes in the text or noisy OCR. A Jaccard similarity score will always be in a range between 0 (complete dissimilarity) and 1 (complete similarity).

@@

[^31]: We have preferred to use the term "digital history" when referring to our own work, in part because the term "digital humanities" has largely come to refer to the work of digital literary and digital media scholars, but primarily because we wish to see digital scholars make disciplinary, rather than interdisciplinary contributions. On the role of disciplines and the importance of field specific argumentation, see, Stephen Robertson, "The Differences between Digital Humanities and Digital History," {PAGE}--{PAGE}, and Cameron Blevins, "Digital History's Perpetual Future Tense," {PAGE}--{PAGE}, both in *Debates in the Digital Humanities 2016*, ed. Matthew K. Gold and Lauren F. Klein (Minneapolis: University of Minnesota Press, 2016). Nevertheless, digital text analysis in the humanities has mostly been published by literary scholars, including Stephen Ramsay, *Reading Machines: Toward an Algorithmic Criticism* (Urbana: University of Illinois Press, 2011); Franco Moretti, *Distant Reading* (New York: Verso, 2013); Matthew L. Jockers, *Macroanalysis: Digital Methods and Literary History* (Urbana: University of Illinois Press, 2013); Ted Underwood, *Why Literary Periods Mattered: Historical Contrast and the Prestige of English Studies* (Stanford: Stanford University Press, 2013); Ryan Cordell and David A. Smith, *Viral Texts: Mapping Networks of Reprinting in 19th-Century Newspapers and Magazines*, NULab for Texts Maps and Networks, Northeastern University (2012--): <http://viraltexts.org>. Related but less well regarded by humanities scholars is work in "culturomics": see Jean-Baptiste Michel et al., "Quantitative Analysis of Culture Using Millions of Digitized Books", *Science* 331, no. 6014 (2011), 176--182; doi:10.1126/science.1199644. On the uselessness of the term "big data," see Ted Underwood, "Against (Talking About) 'Big Data,'" *The Stone and the Shell*, blog post, 10 May 2013: <https://tedunderwood.com/2013/05/10/why-it-matters-that-we-dont-know-what-we-mean-by-big-data/>. For an example of curating a corpus aimed at research questions, see Ted Underwood, Boris Capitanu, Peter Organisciak, Sayan Bhattacharyya, Loretta Auvil, Colleen Fallaw, J. Stephen Downie, "Word Frequencies in English-Language Literature, 1700-1922," dataset, v0.2 (HathiTrust Research Center, 2015) doi:10.13012/J8JW8BSJ.

    Only a few scholars have turned their attention to the computer analysis of legal texts for historical purposes, including Paul Craven, "Detección automática y visualización de dominios específicos similares en documentos: análisis DWIC y su aplicación en el Proyecto Master & Servant \[Automatic Detection and Visualization of Domain-Specific Similarities in Documents: DWIC Analysis and its Application in the Master & Servant Project\]," published on CD-ROM in F. J. A Perez et al., eds., *La Historia en una nueva frontera \[History in a New Frontier\]* (Digibis: Ediciones de la Universidad de Castilla-La Mancha, 1998); Paul Craven and Douglas Hay, "Computer Applications in Comparative Historical Research: The Master & Servant Project at York University, Canada," *History and Computing* 7, no. 2 (1995); Paul Craven and W. Traves, "A General-Purpose Hierarchical Coding Engine and Its Application to Comparative Analysis of Statutes," *Literary and Linguistic Computing* 8, no. 1 (1993): 27--32, doi:10.1093/llc/8.1.27; Eric C. Nystrom and David S. Tanenhaus, "The Future of Digital Legal History: No Magic, No Silver Bullets," *American Journal of Legal History* 56, no. 1 (2016): 150--67, doi:10.1093/ajlh/njv017; Dan Cohen, Frederick Gibbs, Tim Hitchcock, Geoffrey Rockwell, et al., "Data Mining with Criminal Intent," white paper, 31 August 2011, <http://criminalintent.org>.

[^32]: {TODO: list exceptions to Google Books}.

[^33]: In each instance we downloaded an entire volume of sessions laws, statutory compilations, or single-volume codes of procedure and then cropped out irrelevant pages, marginalia and footnoted commentary, leaving only the statutory text. After several trials of various implementations of Tesseract (open source) and I.R.I.S. (proprietary) OCR programs, we determined that Nitro Pro PDF, which relies on I.R.I.S. software, offered the best OCR tool for this project. I.R.I.S. provides slightly more accurate readings of nineteenth-century typefaces than Tesseract, and Nitro Pro's implementation makes words, not characters, the fundamental unit of output. The latter feature made cropping between marginalia and the statute more reliable. We removed hyphenated line breaks and standardized spelling for common terms that evolved over the nineteenth century (e.g., *indorsement*).

[^34]: {TODO: citation on Stephen Johnson Field}
